TODO
- LSEP in Appendix
- Reference in Appendix B
- GitHub Test framework in README and close all tickets
- Read through
- Submit DRAFT
    - Consent Zip
    - EPA form
    - Make them collaborators on GitHub and share repo
    -

\chapter{Self-Appraisal}
\label{app:self_appraisal}

% =============================================================================
\section{Critical Self-Evaluation}
\label{sec:self_eval}
% =============================================================================

This project required simultaneous competence across four distinct disciplines: Unity HDRP development, optical sensor integration, experimental HCI methodology, and statistical analysis. Reflecting critically, the engineering objectives were met to a professional standard while the experimental objectives were partially constrained by hardware factors outside my control.

\subsubsection{What went well}
The formal DFSM architecture was the single best design decision of the project. By enforcing totality across all 36 input-state combinations and routing both modalities through the abstract \texttt{BaseInteractionProvider} contract, I eliminated an entire category of undefined runtime behaviour that had plagued the prototype. The TDD strategy --- writing 29 EditMode assertions and 9 PlayMode integration behaviours before human trials --- provided genuine engineering confidence rather than post-hoc assurance. The telemetry pipeline performed without a single data-loss event across all conducted trials: every task completion exported a well-formed CSV atomically, and the Welford accumulator produced numerically stable grip variance across the full 90\,Hz sample rate. The CONSORT-compliant trial protocol was well-executed; the condition order justification, exclusion criteria and statistical analysis plan were all pre-registered in the methodology rather than retrofitted to the results.

\subsubsection{What went wrong and how I adapted}
The most significant failure was hardware-induced sample reduction. Three of eight enrolled participants were excluded due to Varjo XR-4 subsystem faults producing malformed telemetry exports. This reduced the analysed sample to $n = 5$, below the $n \geq 8$ minimum I had planned for, and meant the Wilcoxon test could not produce $p < .05$ for any two-tailed comparison regardless of effect magnitude. I adapted by foregrounding effect sizes as the primary evidential metric, following the dual-reporting approach of \citet{Rantamaa2023ComparisonReality}, and reframing the study explicitly as a pilot feasibility trial rather than a confirmatory one --- a reframing that is methodologically honest rather than defensive. Had I budgeted two additional enrolment sessions as a contingency for equipment failures, the final sample would have been adequate.

The fixed CTR-first condition order was a deliberate design choice justified by asymmetric prior experience across participants, but in retrospect a partially counterbalanced design --- even with only eight participants --- would have been preferable. I could not implement full counterbalancing without doubling recruitment time, but a blocked crossover assigning four participants to each order would have been feasible within the project timeline.

An early HDRP rendering issue caused a two-week delay at the start of Milestone~2: the Unity HDRP pipeline requires explicit configuration of shadow atlases and volume profiles that are not pre-configured in the default template, and the Varjo XR-4's foveated rendering required additional HDRP compatibility layers that were undocumented at the time. I resolved this by isolating the rendering configuration from the interaction logic through the component architecture, so subsequent changes to either did not require rebuilding the other. The lesson was to validate the full hardware-software stack on a minimal scene before committing to the production architecture --- a principle I now treat as a mandatory first sprint task for any hardware-dependent project.

% =============================================================================
\section{Personal Reflection and Lessons Learned}
\label{sec:reflection}
% =============================================================================

\subsubsection{Technical growth}
Before this project I had not worked with quaternion mathematics in a production context, had not implemented a formal automaton in software, and had not run a within-subjects experimental study. I leave the project confident in all three. The most intellectually satisfying moment was deriving the formal proof that the DFSM accepts exactly the language of valid interaction sequences --- connecting the theoretical computer science of Milestone~1 to the physical interaction behaviour of Milestone~3 in a way that felt genuinely rigorous rather than decorative.

\subsubsection{Project management}
The sprint structure aligned to Milestones 2--4 was effective in preventing scope creep: the Kanban work-in-progress limit of three concurrent issues forced prioritisation decisions that I would otherwise have deferred. The most important lesson was that hardware-dependent research projects require a risk buffer of at least one sprint for device integration failures. I allocated approximately 20\% of Milestone~2 to resolving the HDRP issue, which compressed the gesture recognition development into a tighter window than planned. In future, I would schedule a dedicated hardware validation sprint before the first implementation sprint rather than treating integration as a background task.

\subsubsection{Research methodology}
Running a human study revealed a gap in my preparation: I had designed the statistical analysis plan carefully, but had not stress-tested the trial protocol with a pilot participant before enrolling the main cohort. A single informal dry-run would have identified the headset weight discomfort reported by VR-naïve participants (P04) and allowed me to add a five-minute acclimatisation period to the protocol. This is now the first item I would add to any future experimental checklist.

\subsubsection{Communication}
Writing this dissertation required me to translate between three audiences simultaneously: the non-specialist reader of the Introduction, the methodologist reading the statistics sections, and the software engineer reading the implementation. The discipline of maintaining those distinct registers --- and of resisting the temptation to front-load technical acronyms before a non-expert reader is ready for them --- is a transferable skill I will carry into professional technical writing.

% =============================================================================
\section{Legal, Social, Ethical and Professional Issues}
\label{sec:lsep_issues}
% =============================================================================

\subsection{Legal Issues}
\label{sec:legal_issues}

\subsubsection{Data protection}
The study collected demographic data (age, sex, handedness, VR experience) and task performance data from human participants. All data was handled in accordance with the UK General Data Protection Regulation (UK GDPR) and the University of Leeds data governance policy. Participant data was pseudonymised at the point of collection using the identifiers P01--P08; no name, email address or other directly identifying information was stored alongside the telemetry or questionnaire records. Collected CSV files and questionnaire responses were stored on an encrypted university-managed drive and were not transmitted to third-party services. Contact data used for recruitment was deleted upon study completion.

\subsubsection{Intellectual property and software licensing}
All third-party software components used in this project were operated under their respective licences. The Ultraleap Hyperion SDK is licensed under the Ultraleap SDK Agreement, which permits use in research and educational projects. The Varjo Base SDK is provided under the Varjo Developer Agreement. Unity 2022 LTS was used under a Unity Student licence. NUnit is distributed under the MIT licence. All pre-commit hooks used (\texttt{actionlint}, \texttt{commitlint}, \texttt{pre-commit-hooks}) are open-source tools with MIT or Apache~2.0 licences. The liver mesh \texttt{OBJ} file was sourced from an anonymised clinical CT reconstruction; its provenance was verified by the project supervisor in accordance with the University's research data governance procedures. No proprietary or commercially restricted datasets were used.

\subsubsection{University intellectual property}
Under the University of Leeds intellectual property policy, software developed as part of an undergraduate project remains the intellectual property of the student unless explicitly assigned otherwise. No external stakeholder agreements were in place that would alter this default position.

\subsection{Social Issues}
\label{sec:social_issues}

\subsubsection{Impact on surgical training and planning}
The motivation for this project is explicitly socially beneficial: improving the usability of surgical planning tools has the potential to reduce planning errors, improve resection outcomes for liver cancer patients, and lower the cognitive burden placed on surgeons during high-stakes procedures. However, I acknowledge that a system evaluated on a university convenience sample cannot yet make clinical claims, and that any deployment pathway would require validation on surgical professionals and regulatory evaluation before influencing clinical practice.

\subsubsection{Access and equity}
The hardware required to replicate this study --- a Varjo XR-4 headset ($\approx$\pounds5,000), an Ultraleap LMC2, and a workstation with an NVIDIA RTX~3080 --- is not accessible to most institutions, particularly those in lower-income healthcare systems. This creates a risk that the benefits of high-fidelity mixed-reality surgical planning tools will accrue disproportionately to well-resourced hospitals. Future work should evaluate whether consumer-grade hardware (e.g., Meta Quest with hand tracking) can replicate the findings at a fraction of the cost, broadening the potential reach of the intervention.

\subsubsection{Participant diversity}
The convenience sample was predominantly non-surgical (62.5\%) and young (mean age 22.9 years). The SUS and RTLX results may not generalise to practising surgeons, who bring different spatial reasoning skills, tool-use habits and fatigue profiles. This is acknowledged as a limitation in the main text and motivates the recruitment of domain-specific participants in the confirmatory trial.

\subsubsection{Environmental impact}
Sustained use of a high-performance GPU workstation and the Varjo XR-4 carries an energy cost. Each experimental session consumed approximately 90 minutes of full GPU load. Across eight sessions this represents a modest but non-zero carbon footprint. For a confirmatory trial with $n \geq 20$, I would recommend scheduling sessions to maximise hardware utilisation per day and investigating whether the workstation can be configured to run on renewable energy tariffs.

\subsection{Ethical Issues}
\label{sec:ethical_issues}

This project falls under two University of Leeds block ethical approval categories: 3.2 (Medical Images) and 3.4 (User Testing).

\subsubsection{Medical images}
The liver mesh used as the manipulation target was sourced from a fully anonymised CT reconstruction. No patient-identifiable metadata was present in the \texttt{OBJ} file or associated contour data. The supervisor confirmed the legitimacy of the source prior to use, consistent with the block approval requirement under category~3.2.

\subsubsection{Informed consent and voluntary participation}
Each participant received a written Project Information Sheet prior to enrolment, describing the study purpose, the data to be collected, how data would be used, and the right to withdraw at any time without consequence. Written consent was obtained from all eight enrolled participants before any data collection began. Participation was voluntary; no inducements were offered. Signed consent forms were uploaded to Minerva alongside this report.

\subsubsection{Health and safety}
Participants were briefed on VR-specific risks before headset application, including the possibility of motion sickness, visual discomfort from extended headset use, and the need to remain seated during the session. The inclusion criterion excluded participants with a history of severe motion sickness. No participant reported motion sickness during the study; two reported minor eye fatigue attributable to headset weight after extended sessions, which is consistent with normal use of the Varjo XR-4 for VR-naïve users.

\subsubsection{Data minimisation}
The questionnaires collected only the minimum demographic data required to characterise the sample: age, sex, handedness, VR experience level and modality-specific prior exposure. No medical history, financial information or other sensitive personal data was collected. All responses were anonymised at the point of recording using participant identifiers.

\subsubsection{Right to withdraw}
Three participants were excluded from the analysis due to equipment failure producing malformed telemetry. This exclusion was applied on technical data-integrity grounds and was pre-specified in the exclusion criteria defined before data collection; it did not constitute withdrawal, and all three participants completed the full protocol. No participant exercised the right to withdraw.

\subsection{Professional Issues}
\label{sec:professional_issues}

\subsubsection{BCS Code of Conduct}
This project was conducted in accordance with the BCS Code of Conduct. The simulation was developed with honesty and integrity: known limitations (optical occlusion, fixed condition order, $n = 5$) are reported transparently rather than minimised. All third-party contributions --- SDKs, libraries, pre-commit tools, the liver mesh --- are fully attributed in Appendix~\ref{app:external}. No claims are made in the report that exceed what the evidence supports.

\subsubsection{Reproducibility and open science}
The full source code, pre-commit configuration, GitHub Actions workflow, and anonymised telemetry CSVs are archived in the project repository, referenced in Appendix~\ref{app:external}. The statistical analysis Jupyter notebook is self-contained and reproducible from the raw CSV exports without manual data preparation. This level of reproducibility is consistent with the open science standards increasingly required in HCI and medical informatics research \citep{Nosek2015PromotingScience}.

\subsubsection{Clinical responsibility}
The most significant professional concern with this class of system is the risk of premature clinical adoption. The findings of this pilot study are explicitly framed as hypothesis-generating rather than confirmatory; the main text states directly that the results must not be interpreted as evidence of clinical equivalence between modalities. Any deployment pathway for HTR-based surgical planning tools would require regulatory evaluation (e.g., under the UK Medical Devices Regulations 2002 or the EU MDR 2017/745 for software as a medical device), independent clinical validation, and surgeon-specific training protocols. Communicating these requirements clearly is a professional obligation, and this report fulfils it.

\subsubsection{Software quality and maintainability}
The production codebase follows SOLID design principles: the abstract provider interface satisfies the open-closed principle; the FSM dispatcher satisfies the single-responsibility principle; the telemetry logger is decoupled from both. The 38 automated test assertions provide a regression safety net for any future development. This standard of engineering practice reflects the expectation that research software may be extended by other researchers, and that code submitted as part of a scientific study should be auditable.